Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2019-06-12 papers/lime_diagnostics/visual-diagnostics-of-a-model-explainer.Rnw
more work on paper

\hh{In the field of statistics, there are two main uses for models: interpretation and prediction.} \kg{Machine learning models are often used for the latter purpose. While these models have been proven to provide accurate predictions in a wide range of problems, this predictive ability comes at the cost of interpretability.} These models are typically complex and lack a functional form, so the ability to directly interpret the models is lost. \kg{This has earned machine learning models the reputation as being black boxes. The ability to interpret a model helps to diagnose the model and understand the predictions.} In some areas of application, the need to be able to explain the results from a model is critical. \kg{(These last few sentences could use some work so that the ideas are better connected from one sentence to the next.)} As a result, an area of research on how to explain the output from machine learning algorithms has emerged.

@@

One approach to remedy the problem of uninterpretable models is to develop a model explainer that provides insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. Recently, a handful of model explainers have been developed (reference). LIME (local interpretable model explanations) is one such model explainer (reference).

@@

@@

Predictive models are used in both regression and classification settings. For this paper, we will focus on the classification setting, and in particular, we will only discuss the case with a dichotomous response variable. \hh{Accepted best practice for all}  predictive models is \hh{that during the model fitting process}, the data is divided into training and testing portions. The training data is used to fit the complex model, and the testing data is used to assess the model. \hh{This is done to prevent overfitting and to get a more accurate assessment of the prediction error of the model.} The complex model is then applied to the features in the testing data, and the resulting predictions are compared to ground truth. Of \hh{particular} interest during this assessment are the cases when the model is wrong. In the dichotomous response classification case, there are two ways in which the model can be wrong. The model can make a type I error in which..., or the model can make a type II error in which...

@@

>>>>

\kg{Added in Heike's suggestions and made a few changes of my own to this paragraph:} In the field of statistics, there are two main uses for models: interpretation and prediction. Machine learning models are often used for the latter purpose. While these models have been proven to provide accurate predictions in a wide range of problems, this predictive ability comes at the cost of interpretability. These models are typically complex and lack a functional form, which makes it not possible to directly interpret the models and has earned them the reputation of being black boxes. The ability to interpret a model helps to diagnose the model and understand the predictions, and in some areas of application, the need to be able to explain the results from a model is critical.

@@

\kg{Moved the last sentence from the first paragraph here:} As a result of the desire to be able to explain the predictions, an area of research has emerged on machine learning explain the output from machine learning algorithms. One approach that is being taken to remedy the problem of uninterpretable models is to develop a model explainer that provides insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. Recently, a handful of model explainers have been developed (reference). LIME (local interpretable model explanations) is one such model explainer (reference).

@@

\kg{I moved this paragraph here, because I think this is where I was to first introdude this concept, but it need a lot of reworking to fit in with the intro, and I need to finish writing my thoughts.} Predictive models are used in both regression and classification settings. For this paper, we will focus on the classification setting, and in particular, we will only discuss the case with a dichotomous response variable. \hh{Accepted best practice for all}  predictive models is \hh{that during the model fitting process}, the data is divided into training and testing portions. The training data is used to fit the complex model, and the testing data is used to assess the model. \hh{This is done to prevent overfitting and to get a more accurate assessment of the prediction error of the model.} The complex model is then applied to the features in the testing data, and the resulting predictions are compared to ground truth. Of \hh{particular} interest during this assessment are the cases when the model is wrong. In the dichotomous response classification case, there are two ways in which the model can be wrong. The model can make a type I error in which..., or the model can make a type II error in which...

@@

\subsection{Model Explainers}

@@

\subsection{Iris Data (?)}

\subsection{Bullet Matching Data}
